# Annex. 2 Graphics {.unnumbered}

```{r}
#| echo: false
#| warning: false
#| label: fig-tend_pos
#| fig-alt: >
#|   A graph showing an uptrend and heteroscedastic.
#| cap-location: top
#| fig-cap: >
#|   Bullish and heteroscedastic trend
#| fig-height: 2.5

IBEX |>
  filter(Date > "1995-01" & Date < "1997-01") |>
  mutate(Date = ym(Date)) |>
  ggplot(aes(x = Date, group = 1)) +
  geom_line(aes(y = CloseI), color = "black") +
  theme(axis.text.x = element_text(angle = 90)) +
  ylab("Closing Price") +
  xlab("Date")
```

::: figure-caption
Own elaboration: By using RStudio with the IBEX historical database, obtained from https://finance.yahoo.com/, in the period between 01-1995 to 01-1997.
:::

```{r}
#| echo: false
#| warning: false
#| label: fig-tend_neg
#| fig-alt: >
#|   A graph showing a heteroscedastic downtrend.
#| cap-location: top
#| fig-cap: >
#|   Bearish and heteroscedastic trend
#| fig-height: 2.5

IBEX |>
  filter(Date > "2000-01" & Date < "2003-01") |>
  mutate(Date = ym(Date)) |>
  ggplot(aes(x = Date, group = 1)) +
  geom_line(aes(y = CloseI), color = "black") +
  theme(axis.text.x = element_text(angle = 90)) +
  ylab("Closing Price") +
  xlab("Date")
```

::: figure-caption
Own elaboration: Through the use of RStudio with the IBEX historical database, obtained from https://finance.yahoo.com/, in the period between 01-2000 to 01-2003.
:::

```{r}
#| echo: false
#| warning: false
#| label: fig-tend_nul
#| fig-alt: >
#|   A graph showing a heteroscedastic downtrend.
#| cap-location: top
#| fig-cap: >
#|   No trend, homoscedastic and stationary
#| fig-height: 2.5

IBEX |>
  mutate(CloseI = Delt(CloseI)) |>
  na.omit() |>
  filter(Date > "2000-01" & Date < "2003-01") |>
  mutate(Date = ym(Date)) |>
  ggplot(aes(x = Date, group = 1)) +
  geom_line(aes(y = CloseI), color = "black") +
  theme(axis.text.x = element_text(angle = 90)) +
  ylab("Closing Profitability") +
  xlab("Date")
```

::: figure-caption
Own elaboration: By using RStudio with the IBEX historical database obtained from https://finance.yahoo.com/, in the period from 01-2000 to 01-2003, using the returns calculated from the closing price.
:::

```{r}
#| echo: false
#| warning: false
#| label: fig-est_err
#| fig-alt: >
#|   A graph showing the seasonality and error components.
#| fig-cap: >
#|   Decomposition: seasonality and error
#| fig-height: 2.5

IBEXts <- IBEX |>
  filter(Date > "2000-01" & Date < "2023-01") |>
  select(CloseI) |>
  ts(frequency = 12) |>
  decompose()

pp1 <- IBEX |>
  filter(Date > "2000-01" & Date < "2023-01") |>
  mutate(seasonal = as.vector(IBEXts$seasonal)) |>
  mutate(Date = ym(Date)) |>
  ggplot(aes(x = Date, group = 1)) +
  geom_line(aes(y = seasonal), color = "black") +
  theme(axis.title.x = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  ylab("Seasonality") +
  xlab("Date")
pp2 <- IBEX |>
  filter(Date > "2000-01" & Date < "2023-01") |>
  mutate(Random = as.vector(IBEXts$random)) |>
  mutate(Date = ym(Date)) |>
  ggplot(aes(x = Date, group = 1)) +
  geom_line(aes(y = Random), color = "black") +
  theme(axis.text.x = element_text(angle = 90)) +
  ylab("Residual") +
  xlab("Date")
grid.arrange(pp1, pp2, ncol = 1, heights = c(1.4, 2))
```

::: figure-caption
Own elaboration: By using RStudio with the IBEX historical database obtained from https://finance.yahoo.com/, decomposing the time series made up of the observations that cover the period from 01-2000 to 01-2023.
:::

```{r}
#| echo: false
#| warning: false
#| label: fig-acf_pacf
#| fig-alt: >
#|   Autocorrelation and partial correlation correlograms.
#| fig-cap: >
#|   Correlogram
#| fig-height: 2.5

ts <- IBEX |>
  filter(Date > "2000-01" & Date < "2023-01") |>
  select(CloseI) |>
  ts(frequency = 12)
grid.arrange(ggAcf(ts), ggPacf(ts), ncol = 2)
```

::: figure-caption
Own elaboration: By using RStudio.
:::

```{r}
#| echo: false
#| warning: false
#| label: fig-tend_const_nxt
#| fig-alt: >
#|   A graph that shows the inconsistency in the registration 
#|   of the adjusted closing price of the series corresponding 
#|   to the company NUEVA EXPRESION TEXTIL, S.A., when 
#|   observing that there is no variation of the adjusted closing 
#|   price in the period corresponding to 06-2009 to 04-2014.
#| cap-location: top
#| fig-cap: >
#|   Constant trend in the adjusted closing prices of 
#|   the company "Nueva Expresión Textil S.A"
#| fig-height: 4 

fig_df <- select_emps[["NXT.MC"]] |>
  mutate(Date = as.Date(Date))
ggplot(fig_df, aes(x=Date, y=Adjusted, group = 1)) +
  geom_line(color="#065AD8") +
  annotate(geom = 'point',x=fig_df$Date[142],y=2.6,size=26, shape=1, color="red")
```

::: figure-caption
Own elaboration: Based on the data obtained from @yahoofin corresponding to the company "Nueva Expreción Textil S.A" in the period from January 31, 2000 to February 28, 2023.
:::

```{r}
#| echo: false
#| warning: false
#| label: fig-radi_chang_btc
#| fig-alt: >
#|   A chart showing sharp changes in adjusted closing 
#|   prices relative to the rest of the values in the 
#|   series, which may indicate a miscalculation 
#|   of values over a period of time.
#| cap-location: top
#| fig-cap: >
#|   Sudden price changes that reflect an erroneous calculation of adjusted closing prices, "BANKINTER,S.A."
#| fig-height: 4

fig_df <- select_emps[["BKT.MC"]] |>
  mutate(Date = as.Date(Date))
ggplot(fig_df, aes(x=Date, y=Adjusted, group = 1)) +
  geom_line(color="#065AD8") +
  annotate(geom = 'point',x=fig_df$Date[1],y=15,size=76, shape=1, color="red")
```

::: figure-caption
Own elaboration: Based on the data obtained from @yahoofin corresponding to the company "Nueva Expreción Textil S.A" in the period from January 31, 2000 to February 28, 2023.
:::

```{r}
#| echo: false
#| warning: false
#| label: fig-ind_evo_oo
#| cap-location: top
#| fig-cap: >
#|   Evolution of indicators – Entries with one observation

data <- do.call(cbind,resultscnnlstmssing1$resultsD)
data |>
  rowwise() |>
  mutate(
    Date = `001.Date`,
    meanmse = mean(c_across(contains("mse"))),
    meanrsqrd = mean(c_across(contains("rsqrd")))
    ) |>
  select(
    Date, meanmse,meanrsqrd
  )|>
  mutate(
    Date = as.Date(Date)) |>
  ggplot(aes(x = Date, group = 1)) +
  geom_line(aes(y = meanmse, color = "MSE")) +
  geom_line(aes(y = meanrsqrd, color = "R2")) +
  scale_color_manual(values = c("blue", "green")) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x = "Date", y = "Values", color = "Indicators")
```

::: figure-caption
Own elaboration: Through the use of R and Rstudio.
:::

```{r}
#| echo: false
#| warning: false
#| label: fig-ind_evo_twoo
#| cap-location: top
#| fig-cap: >
#|   Evolution of the indicators – Entries with two observations

data <- do.call(cbind,resultscnnlstmssing2$resultsD)
data |>
  rowwise() |>
  mutate(
    Date = `001.Date`,
    meanmse = mean(c_across(contains("mse"))),
    meanrsqrd = mean(c_across(contains("rsqrd")))
    ) |>
  select(
    Date, meanmse,meanrsqrd
  )|>
  mutate(
    Date = as.Date(Date)) |>
  ggplot(aes(x = Date, group = 1)) +
  geom_line(aes(y = meanmse, color = "MSE")) +
  geom_line(aes(y = meanrsqrd, color = "R2")) +
  scale_color_manual(values = c("blue", "green")) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x = "Date", y = "Values", color = "Indicators")
```

::: figure-caption
Own elaboration: Through the use of R and Rstudio.
:::

```{r}
#| echo: false
#| warning: false
#| label: fig-ind_evo_threeo
#| cap-location: top
#| fig-cap: >
#|   Evolution of the indicators – Entries with three observations

data <- do.call(cbind,resultscnnlstmssing3$resultsD)
data |>
  rowwise() |>
  mutate(
    Date = `001.Date`,
    meanmse = mean(c_across(contains("mse"))),
    meanrsqrd = mean(c_across(contains("rsqrd")))
    ) |>
  select(
    Date, meanmse,meanrsqrd
  )|>
  mutate(
    Date = as.Date(Date)) |>
  ggplot(aes(x = Date, group = 1)) +
  geom_line(aes(y = meanmse, color = "MSE")) +
  geom_line(aes(y = meanrsqrd, color = "R2")) +
  scale_color_manual(values = c("blue", "green")) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x = "Date", y = "Values", color = "Indicators")
```

::: figure-caption
Own elaboration: Through the use of R and Rstudio.
:::

```{r}
#| echo: false
#| warning: false
#| label: fig-pf_evo_oo
#| cap-location: top
#| fig-cap: >
#|   Evolution of portfolios and the IBEX – Entries with one observation

data <- do.call(cbind,resultscnnlstmssing1$resultsD)
data_summary <- data |>
  mutate(
    Date = `001.Date`,
    IBEX = `001.IBEX`,
    Means = resultscnnlstmssing1$resultsMEANS) |>
  mutate_at(vars(contains("Portre")), ~ cumsum(.)) |>
  mutate(
    IBEX = cumsum(IBEX),
    Means = cumsum(Means)) |>
  group_by(Date) |>
  summarize(
    meanPortre = mean(c_across(contains("Portre"))),
    max_y = max(c_across(contains("Portre"))),
    min_y = min(c_across(contains("Portre"))),
    min_5 = unname(quantile(c_across(contains("Portre")),0.05)),
    max_95 = unname(quantile(c_across(contains("Portre")),0.95)),
    IBEX = IBEX,
    Means = Means)
data_summary |>
  mutate(
    Date = as.Date(Date)) |>
ggplot(aes(x = Date)) +
  geom_ribbon(aes(ymin = min_y, ymax = min_5), fill = "blue", alpha = 0.3) +
  geom_ribbon(aes(ymin = max_y, ymax = max_95), fill = "blue", alpha = 0.3) +
  geom_ribbon(aes(ymin = min_5, ymax = max_95), fill = "blue", alpha = 0.6) +
  geom_line(
    aes(y = meanPortre, color = "Mean RNA1"),
    linetype = "dashed") +
  geom_line(aes(y = max_y), color = "blue") +
  geom_line(aes(y = min_y), color = "blue") +
  geom_line(aes(y = max_95), color = "blue") +
  geom_line(aes(y = min_5, color = "RNA1")) +
  geom_line(aes(y = IBEX, color = "IBEX")) +
  geom_line(aes(y = Means, color = "Means")) +
  scale_color_manual(
    values = c(
      "Mean RNA1"="blue",
      "RNA1" = "blue",
      "IBEX" = "red",
      "Means" = "green")) +
  guides(
    color = guide_legend(
      override.aes = list(
        linetype = c("solid","dashed","solid","solid"))))+
  labs(x = "Date",
       y = "Values",
       color = "Legend")+
  theme_minimal()
```

::: figure-caption
Own elaboration: Through the use of R and Rstudio.
:::

```{r}
#| echo: false
#| warning: false
#| label: fig-pf_evo_twoo
#| cap-location: top
#| fig-cap: >
#|   Evolution of the portfolios and the IBEX – Entries with two observations

data <- do.call(cbind,resultscnnlstmssing2$resultsD)
data_summary <- data |>
  mutate(
    Date = `001.Date`,
    IBEX = `001.IBEX`,
    Means = resultscnnlstmssing2$resultsMEANS) |>
  mutate_at(vars(contains("Portre")), ~ cumsum(.)) |>
  mutate(
    IBEX = cumsum(IBEX),
    Means = cumsum(Means)) |>
  group_by(Date) |>
  summarize(
    meanPortre = mean(c_across(contains("Portre"))),
    max_y = max(c_across(contains("Portre"))),
    min_y = min(c_across(contains("Portre"))),
    min_5 = unname(quantile(c_across(contains("Portre")),0.05)),
    max_95 = unname(quantile(c_across(contains("Portre")),0.95)),
    IBEX = IBEX,
    Means = Means)
data_summary |>
  mutate(
    Date = as.Date(Date)) |>
ggplot(aes(x = Date)) +
  geom_ribbon(aes(ymin = min_y, ymax = min_5), fill = "blue", alpha = 0.3) +
  geom_ribbon(aes(ymin = max_y, ymax = max_95), fill = "blue", alpha = 0.3) +
  geom_ribbon(aes(ymin = min_5, ymax = max_95), fill = "blue", alpha = 0.6) +
  geom_line(
    aes(y = meanPortre, color = "Mean RNA2"),
    linetype = "dashed") +
  geom_line(aes(y = max_y), color = "blue") +
  geom_line(aes(y = min_y), color = "blue") +
  geom_line(aes(y = max_95), color = "blue") +
  geom_line(aes(y = min_5, color = "RNA2")) +
  geom_line(aes(y = IBEX, color = "IBEX")) +
  geom_line(aes(y = Means, color = "Means")) +
  scale_color_manual(
    values = c(
      "Mean RNA2"="blue",
      "RNA2" = "blue",
      "IBEX" = "red",
      "Means" = "green")) +
  guides(
    color = guide_legend(
      override.aes = list(
        linetype = c("solid","dashed","solid","solid"))))+
  labs(x = "Date",
       y = "Values",
       color = "Legend")+
  theme_minimal()
```

::: figure-caption
Own elaboration: Through the use of R and Rstudio.
:::

```{r}
#| echo: false
#| warning: false
#| label: fig-pf_evo_threeo
#| cap-location: top
#| fig-cap: >
#|   Evolution of the portfolios and the IBEX – Entries with three observations

data <- do.call(cbind,resultscnnlstmssing3$resultsD)
data_summary <- data |>
  mutate(
    Date = `001.Date`,
    IBEX = `001.IBEX`,
    Means = resultscnnlstmssing3$resultsMEANS) |>
  mutate_at(vars(contains("Portre")), ~ cumsum(.)) |>
  mutate(
    IBEX = cumsum(IBEX),
    Means = cumsum(Means)) |>
  group_by(Date) |>
  summarize(
    meanPortre = mean(c_across(contains("Portre"))),
    max_y = max(c_across(contains("Portre"))),
    min_y = min(c_across(contains("Portre"))),
    min_5 = unname(quantile(c_across(contains("Portre")),0.05)),
    max_95 = unname(quantile(c_across(contains("Portre")),0.95)),
    IBEX = IBEX,
    Means = Means)
data_summary |>
  mutate(
    Date = as.Date(Date)) |>
ggplot(aes(x = Date)) +
  geom_ribbon(aes(ymin = min_y, ymax = min_5), fill = "blue", alpha = 0.3) +
  geom_ribbon(aes(ymin = max_y, ymax = max_95), fill = "blue", alpha = 0.3) +
  geom_ribbon(aes(ymin = min_5, ymax = max_95), fill = "blue", alpha = 0.6) +
  geom_line(
    aes(y = meanPortre, color = "Mean RNA3"),
    linetype = "dashed") +
  geom_line(aes(y = max_y), color = "blue") +
  geom_line(aes(y = min_y), color = "blue") +
  geom_line(aes(y = max_95), color = "blue") +
  geom_line(aes(y = min_5, color = "RNA3")) +
  geom_line(aes(y = IBEX, color = "IBEX")) +
  geom_line(aes(y = Means, color = "Means")) +
  scale_color_manual(
    values = c(
      "Mean RNA3"="blue",
      "RNA3" = "blue",
      "IBEX" = "red",
      "Means" = "green")) +
  guides(
    color = guide_legend(
      override.aes = list(
        linetype = c("solid","dashed","solid","solid"))))+
  labs(x = "Date",
       y = "Values",
       color = "Legend")+
  theme_minimal()
```

::: figure-caption
Own elaboration: Through the use of R and Rstudio.
:::
